---
title: "Some R Demo: Vectors and Matrices"
author: "STAT 445/645"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## 1. Example: Load a built-in dataset

```{r}
data(iris)
head(iris)
dim(iris)
str(iris)
```

## 2. Vectors

### Example: Creating vectors

```{r}
x <- c(1, 3.3, -0.5)
y <- c(6.7, -1.2, 9)

x
y
length(x)
length(y)
```

### Example: use R as a calculator

```{r}
x - y
2 * x
-1 * x
```

### Inner product and length

```{r}
# inner product (two equivalent ways)
sum(x * y) # run: ?sum, to check the documentation of a R function
t(x) %*% y # run: ?t

# Euclidean length (norm)
sqrt(sum(x^2))
sqrt(as.numeric(t(x) %*% x))
```

## 3. Matrices

### Example: Creating matrices

```{r}
A <- matrix(c(
  1.0,  8.0,
  2.3,  6.0,
  4.7, -9.5
), nrow = 3, byrow = TRUE)

B <- matrix(c(
  2, 1,
  0, 3
), nrow = 2, byrow = TRUE)

C <- diag(c(1,6,3.5))

A
B
C
dim(A)
dim(B)
dim(C)
```

### Transpose, scalar multiplication, addition

```{r}
t(A)
3 * A
A + A
A - A
```

### Matrix multiplication

```{r}
A %*% B

# Element-wise multiplication (not matrix multiplication)
A * A
```

## 4. Matrix–vector multiplication

```{r}
v <- c(1, -2)
A %*% v
```



## 6. Solve a linear system, inverse of a matrix


```{r}
C <- matrix(c(
  2, 1,
  1, 3
), nrow = 2, byrow = TRUE)

I <- diag(c(1,1))

C
I

solve(C) # inverse of matrix C

x_hat <- solve(C, I) # solve Cx = I
x_hat


# Even large matrix
D <- matrix(rnorm(100), nrow = 10, byrow = TRUE)
D
solve(D)
```
## 6. Determinant of a square matrix 


```{r}
C <- matrix(c(3, 1, 6,
              7, 4, 5,
              2, -7, 1), nrow = 3, byrow = TRUE)

det(C)
```



## 7. Eigenvectors, eigenvalues of a square matrix 


```{r}
B <- matrix(rnorm(9), nrow = 3)
A <- t(B) %*% B

e <- eigen(A)   
e$values
e$vectors
```
